Control Aware Radio Resource Allocation in Low Latency Wireless Control Systems

### 2019-02-04 Low_latency_control/response/response_main.tex

\item[{\bf[C3]}] Reviewers raised numerous questions regarding some of the practical considerations in Section VI, such as backtracking parameter selection as well as satisfying the non-stationary assumption in Assumption 4. To address these concerns, we have added additional details to Section VI to clarify many of these points, and an additional remark after Assumption 4. We thank the reviewers for raising these questions.

\end{itemize}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised concerns about the placement of the work within the URLL control and communication framework and literature. The reviewer also brings up points of clarification regarding the problem formulation. Each concern is addressed individually below.

@@

\issue{ The authors need to recast the framework within ultra-reliable and low-latency control and communication.}
\answer{\red{To be written}}

\issue{ It was unclear whether the focus is on communication for a given control or the other way around. Taht is currently it looks like the solution is based on sequential design.}
\answer{We thank the reviewer for brining up this point of clarification. The focus of this work is designing a scheduling protocol for a given system and control. As such, the control is assumed to be given, or chosen agnostic to the communication, while the communication is designed relative to both the system dynamics and choice of control. \blue{We have added a sentence to Section II to clarfiy this point.} }

@@

\issue{ The article would improve a lot if specific use cases are defined.}
\answer{\red{To be written}}

\issue{ Linear control was adopted (for tractability) but clearly the nonlinearity is needed}
\answer{The reviewer brings up a reasonable point about the formulation of the wireless control problem. Indeed, many practical systems feature both nonlinear dynamics and potentially non-linear controls. As stated, the proposed approach focuses on linear systems and controls so that we may derive analytic expressions for the control-aware adaptive packet delivery rates. In practice, nonlinear systems may be linearized for the purposes of obtaining such packet delivery rate targets in (23) and performing the proposed CALLS method as written. \blue{To demonstrate how this may be done, in the revised manuscript we have included additional simulations in Section VI in which we control a non-linear system using the CALLS method to schedule plant communications. In these simulations, while the actual dynamics and control being applied are nonlinear in nature, a linearization of the dynamics is used solely to derive the necessary PDR targets as in expression (23). Here, we demonstrate that the proposed method is applicable even in nonlinear systems.}   }

@@

\issue{Reliability as not addressed however it is intertwined with latency}
\answer{\red{To be written}}

\issue{Authors did not cite recent and fundamental articles within URLLC:}
\answer{We thank the reviewer for pointing out these references. They have been added to the revised manuscript, as well as \blue{ included in an added discussion of URLLC in the introduction (Section I).} }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised concerns about the readability of the abstract and made some suggestions regarding the simulation results. Each concern is addressed individually below:

\issue{ Currently, it is not easy to follow the abstract. It should be revised. The contributions of the this paper should be highlighted.}
\answer{We thank the reviewer for raising this point. \blue{We have revised the abstract in the manuscript to be more clear and the highlight the specific contributions of the paper.}}

@@

\issue{  In simulation part, it would be great if the authors could compare the proposed CALLS method with other methods if possible, since there are also some papers discussing communication resource allocation for multiple control loops.}
\answer{\red{To be written}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised points of clarification regarding the system setup and proposed additional comparisons to be made in the simulation results. Each concern is addressed individually below:

\issue{ Why do you only consider the AWGN channels?  How will your scheme performance when considering fading channels? Is the fading channel also a typical scenario for WiFi based in-door wireless control system?  }
\answer{We thank the reviewer for raising this point, as this aspect of the model is poorly worded in the previous version of the manuscript. We do indeed consider a fading channel to model the wireless channel conditions (specifically, IEEE Channel Model E). In our simulations, we use AWGN curves to evaluate packet error probabilities, but the curves are evaluated at the complete SNR values, which take into account the instantaneous fast fading channel conditions. In addition, we point out that IEEE Channel Model E is the current standard for measuring channel conditions in an indoor environment. More sophisticated models may be developed, but are outside the scope of this work. \blue{We have revised the manuscript to clarify these points regarding the wireless channel model.}}

\issue{In your system, how do you obtain accurate channel state information? E.g. for AWGN channels, you need to test the noise variance. Is your scheme sensitive to channel estimation errors?}
\answer{This is a good point raised by the reviewer. \red{To be written.}}

### 2019-02-05 Low_latency_control/response/response_main.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised concerns about the placement of the work within the URLL control and communication framework and literature. The reviewer also brings up points of clarification regarding the problem formulation. Each concern is addressed individually below.

\issue{ The authors need to recast the framework within ultra-reliable and low-latency control and communication.}
\answer{\red{To be written}}

\issue{ It was unclear whether the focus is on communication for a given control or the other way around. Taht is currently it looks like the solution is based on sequential design.}
\answer{We thank the reviewer for brining up this point of clarification. The focus of this work is designing a scheduling protocol for a given system and control. As such, the control is assumed to be given, or chosen agnostic to the communication, while the communication is designed relative to both the system dynamics and choice of control. We have added a sentence to Section II to clarify this point. }

@@

\issue{Authors did not cite recent and fundamental articles within URLLC:}
\answer{We thank the reviewer for pointing out these references. They have been added to the revised manuscript, as well as \blue{ included in an added discussion of URLLC in the introduction (Section I).} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%   R   E   V   I   E   W   E   R    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\reviewer

We thank the reviewer for the comments, feedback, and questions. The reviewer raised concerns about the readability of the abstract and made some suggestions regarding the simulation results. Each concern is addressed individually below:

\issue{ Currently, it is not easy to follow the abstract. It should be revised. The contributions of the this paper should be highlighted.}
\answer{We thank the reviewer for raising this point. We have revised the abstract in the manuscript to be more clear and the highlight the specific contributions of the paper.}

### 2019-02-05 Low_latency_control/revised/problem_formulation.tex
Add files via upload

%%%%%%%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\input{images/tikz_wncs4}
\caption{Wireless control system with $m$ independent systems. Each system contains a sensor that measure state information, which is transmitted to the controller over a wireless channel. The state information is used by the controller to determine control policies for each of the systems.The communication is assumed to be wireless in the uplink and ideal in the downlink.}
\label{fig_wcs}
\end{figure}

@@

Because the control loop in Figure \ref{fig_wcs} is closed over a wireless channel, there exists a possibility at each cycle $k$ that the transmission fails and state information is not received by the controller. We refer to this as the ``open-loop' configuration; when state information is received, the system operates in ``closed-loop.'' As such, it is necessary to define the system dynamics in both configurations. Consider a generic linear control, in which the input being determined as $ \bbu_{i,k} = \bbK_i \bbx_{i,k}$ for some matrix $\bbK_i \in \reals^{q \times p}$. Many common control policies indeed can be formulated in such a manner, such as LQR control. In general, this matrix $\bbK$ is chosen such as that the closed loop dynamic matrix $\bbA + \bbB \bbK$ is stable, i.e. has all eigenvalues less that 1. Thus, application of this control over time will drive the state $\bbx_{i,k} \rightarrow 0$ as $k \rightarrow \infty$. As the controller does not always have access to state information, we alternatively consider the estimate of state information of device $i$ known to the controller at time $k$ as
%
\begin{align}\label{eq_state_est}
\hbx^{(l_i)}_{i,k} :=(\bbA_i + \bbB_i\bbK_i)^{l_i} \bbx_{i,k-l_i},
\end{align}
%
where $k-l_i \geq k-1$ is the last time instance in which control system $i$ was closed. There are two important things to note in \eqref{eq_state_est}. First, this is the estimated state \emph{before} a transmission has been attempted at time $k$; hence, $l_i = 1$ when state information was received at the previous time. Second, observe that in \eqref{eq_state_est} we assume that the AP/controller has knowledge of the dynamics $\bbA_i$ and $\bbB_i$, as well as the linear control matrix $\bbK_i$. Any gap in this knowledge of dynamics is captured in the noise $\bbw_k$ in the actual dynamics in \eqref{eq_control_orig}. Note that the estimated state \eqref{eq_state_est} is used in place of the true state in both the determination of the control \emph{and} the radio resource allocation decisions as discussed later in this paper.

@@

%%%%%%% F I G U R E %%%%%%%%%%%
\begin{figure}
%\centering
%\input{multiplex.tex}
\includegraphics[width=0.5\textwidth]{images/schedule2.png}
\caption{Multiplexing of frequencies (RU) and time (PPDU) in IEEE 802.11ax transmission window (formally referred as Transmission Opportunity or TXOP in the standard. The total transmission time is the time of all PPDUs, including the overhead of trigger frames (TF) and acknowledgments.}
\label{fig_multiplex}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%% T A B L E %%%%%%%%%%%
\begin{table}
\input{images/mcs.tex}
\caption{Data rates for MCS configurations in IEEE 802.11ax for 20MHz channel. The modulation type and coding rate in the first 2 columns together specify a PDR function $q(\bbmu,\bbsigma)$ for RU $\bbsigma$. The data rate in the third column specifies the associated transmission time $\tau(\mu, \bbsigma)$. }
\label{tab_mcs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%% F I G U R E %%%%%%%%%%%%%%%%5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\input{../images/tikz_wncs4}
\caption{Wireless control system with $m$ independent systems. Each system contains a sensor that measure state information, which is transmitted to the controller over a wireless channel. The state information is used by the controller to determine control policies for each of the systems.The communication is assumed to be wireless in the uplink and ideal in the downlink.}
\label{fig_wcs}
\end{figure}

@@

Because the control loop in Figure \ref{fig_wcs} is closed over a wireless channel, there exists a possibility at each cycle $k$ that the transmission fails and state information is not received by the controller. We refer to this as the ``open-loop' configuration; when state information is received, the system operates in ``closed-loop.'' As such, it is necessary to define the system dynamics in both configurations. Consider a generic linear control, in which the input being determined as $ \bbu_{i,k} = \bbK_i \bbx_{i,k}$ for some matrix $\bbK_i \in \reals^{q \times p}$. Many common control policies indeed can be formulated in such a manner, such as LQR control. In general, this matrix $\bbK$ is chosen such as that the closed loop dynamic matrix $\bbA + \bbB \bbK$ is stable, i.e. has all eigenvalues less that 1. Thus, application of this control over time will drive the state $\bbx_{i,k} \rightarrow 0$ as $k \rightarrow \infty$. \red{We assume that this choice of $\bbK$ is given---in other words, the controller is pre-designed with respect to ideal closed loop behavior.} As the controller does not always have access to state information, we alternatively consider the estimate of state information of device $i$ known to the controller at time $k$ as
%
\begin{align}\label{eq_state_est}
\hbx^{(l_i)}_{i,k} :=(\bbA_i + \bbB_i\bbK_i)^{l_i} \bbx_{i,k-l_i},
\end{align}
%
where $k-l_i \geq k-1$ is the last time instance in which control system $i$ was closed. There are two important things to note in \eqref{eq_state_est}. First, this is the estimated state \emph{before} a transmission has been attempted at time $k$; hence, $l_i = 1$ when state information was received at the previous time. Second, observe that in \eqref{eq_state_est} we assume that the AP/controller has knowledge of the dynamics $\bbA_i$ and $\bbB_i$, as well as the linear control matrix $\bbK_i$. Any gap in this knowledge of dynamics is captured in the noise $\bbw_k$ in the actual dynamics in \eqref{eq_control_orig}. Note that the estimated state \eqref{eq_state_est} is used in place of the true state in both the determination of the control \emph{and} the radio resource allocation decisions as discussed later in this paper.

@@

%%%%%%% F I G U R E %%%%%%%%%%%
\begin{figure}
%\centering
%\input{multiplex.tex}
\includegraphics[width=0.5\textwidth]{../images/schedule2.png}
\caption{Multiplexing of frequencies (RU) and time (PPDU) in IEEE 802.11ax transmission window (formally referred as Transmission Opportunity or TXOP in the standard. The total transmission time is the time of all PPDUs, including the overhead of trigger frames (TF) and acknowledgments.}
\label{fig_multiplex}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%% T A B L E %%%%%%%%%%%
\begin{table}
\input{../images/mcs.tex}
\caption{Data rates for MCS configurations in IEEE 802.11ax for 20MHz channel. The modulation type and coding rate in the first 2 columns together specify a PDR function $q(\bbmu,\bbsigma)$ for RU $\bbsigma$. The data rate in the third column specifies the associated transmission time $\tau(\mu, \bbsigma)$. }
\label{tab_mcs}
\end{table}
%%%%%%%%%%%%%%%%%%%%%

### 2019-02-05 Low_latency_control/revised/root.tex
Add files via upload

\input{mysymbol.sty}
\input{my_sections}
%\input{TikZGallery-master/Sty/TikZGallery-Preamble}
%\renewcommand{\baselinestretch}{.995}
\newcommand{\QED}{\hfill\ensuremath{\blacksquare}}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}

@@

%\renewcommand \blue[1]{}  \renewcommand \red[1]{}

@@

\author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad Konstantinos Gatsis$^*$ \\ \textup{Dave Cavalcanti$^\dagger$\quad Nageen Himayat$^{\dagger}$ \quad Alejandro Ribeiro$^*$}
\thanks{Supported by the Intel Science and Technology Center for Wireless Autonomous Systems and ARL DCIST CRA W911NF-17-2-0181. The authors are with the ($^*$)Department of Electrical and Systems Engineering, University of Pennsylvania and ($^\dagger$)Wireless Communications Research, Intel Corporation. Email: maeisen@seas.upenn.edu, mamun.rashid@intel.com, kgatsis@seas.upenn.edu, dave.cavalcanti@intel.com, nageen.himayat@intel.com, aribeiro@seas.upenn.edu.}}

@@

%
\begin{abstract}
%We consider the problem of allocating radio resources, such as frequency, bandwidth, data rates and time slots, over wireless communication links to control a series of independent wireless control systems. Low-latency transmissions are necessary in control systems that contain high sampling rates, thus necessitating fast transmissions. Compared to wired links, wireless links are highly susceptible to communication errors arising from channel noise and interference. Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates. However, the effect of the communication link errors on the control system performance is not static but depends dynamically on the control system state. Traditional resource allocation methods over wireless links are unaware of control system states and thus are designed with static link error performance, which could be very inefficient if  error performance assumption is too conservative (e.g. too low error rate target), or unacceptably unreliable if the assumption is too relaxed. In this paper, we propose a novel control-communication co-design approach to this low-latency resource allocation problem. We develop methods that incorporate control state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links connecting the different components of the control systems. Control systems that are closer to instability or further from a desired range in a given control cycle are given priority in the scheduling. Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy respective Lyapunov control targets and make scheduling decisions to best meet targets while minimizing total transmission time. The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.
We consider the problem of allocating radio resources over wireless communication links to control a series of independent wireless control systems. Low-latency transmissions are necessary in enabling time-sensitive control systems to operate over wireless links with high reliability. Achieving fast data rates over wireless links thus comes at the cost of reliability in the form of high packet error rates compared to wired links due to channel noise and interference.  However, the effect of the communication link errors on the control system performance depends dynamically on the control system state. We propose a novel control-communication co-design approach to the low-latency resource allocation problem. We incorporate control and channel state information to make scheduling decisions over time on frequency, bandwidth and data rates across the next-generation Wi-Fi based wireless communication links that close the control loops.  Control systems that are closer to instability or further from a desired range in a given control cycle are given higher packet delivery rate targets to meet. Rather than a simple priority ranking, we derive precise packet error rate targets for each system needed to satisfy stability targets and make scheduling decisions to meet such targets while reducing total transmission time. The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.

\end{abstract}
%

@@

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\urlstyle{same}
\bibliographystyle{IEEEtran}
\bibliography{wireless_ll_control,scheduling_control}

>>>>

\input{../mysymbol.sty}
\input{../my_sections}
\input{../TikZGallery-master/Sty/TikZGallery-Preamble}
%\renewcommand{\baselinestretch}{.995}
\newcommand{\QED}{\hfill\ensuremath{\blacksquare}}
\newcommand{\norm}[1]{\ensuremath{\left\| #1 \right\|}}

@@

%\renewcommand \blue[1]{}  \renewcommand \red[1]{

@@

\author{Mark Eisen$^*$ \quad Mohammad M. Rashid$^\dagger$ \quad Konstantinos Gatsis$^*$ \\ \textup{Dave Cavalcanti$^\dagger$\quad Nageen Himayat$^{\dagger}$ \quad Alejandro Ribeiro$^*$}
\thanks{Supported by the Intel Science and Technology Center for Wireless Autonomous Systems. The authors are with the ($^*$)Department of Electrical and Systems Engineering, University of Pennsylvania and ($^\dagger$)Wireless Communications Research, Intel Corporation. Email: maeisen@seas.upenn.edu, mamun.rashid@intel.com, kgatsis@seas.upenn.edu, dave.cavalcanti@intel.com, nageen.himayat@intel.com, aribeiro@seas.upenn.edu.}}

@@

%
\begin{abstract}
\red{We consider the problem of allocating radio resources over wireless communication links to control a series of independent wireless control systems. Low-latency transmissions are necessary in enabling time-sensitive control systems with high sampling rates to operate over wireless links. Enabling low-latency through fast data rates comes at the cost of reliability in the form of higher packet error rates due to channel noise and interference.  However, the cost of such communication link errors on the control system performance depends dynamically on the control system state. We propose a novel control-aware communication design to the low-latency resource allocation problem. In our proposed method, we incorporate both control and channel state information to scheduling transmissions across time slots, frequency bands, and data rates using the next-generation Wi-Fi scheduling architecture.  Control systems that are closer to instability or further from a desired range in a given control cycle are given higher packet delivery rate targets to meet. Rather than a simple priority ranking, we derive precise adaptive packet error rate targets for each system needed to satisfy control-specific performance requirements. We use these adaptive rate targets to make scheduling decisions that reduce total transmission time. The resulting Control-Aware Low Latency Scheduling (CALLS) method is tested in numerous simulation experiments that demonstrate its effectiveness in meeting control-based goals under tight latency constraints relative to control-agnostic scheduling.}
\end{abstract}
%

@@

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\urlstyle{same}
\bibliographystyle{IEEEtran}
\bibliography{../wireless_ll_control,../scheduling_control}
